From Reversible Programs to \\ Univalent Universes and Back

### 2017-06-26 upi.tex
toffoli

\usepackage{bbold}
\usepackage{bussproofs}
\usepackage{keystroke}
\usepackage{comment}
\usepackage{tikz}
\usepackage[inline]{enumitem}

@@

Starting from the physical principle of ``conservation of
information''~\cite{Hey:1999:FCE:304763,fredkin1982conservative}, James and
Sabry~\cite{James:2012:IE:2103656.2103667} propose a family of programming
languages $\Pi$ in which computation preserve information. Technically,
computations are \emph{type isomorphisms} which, at least in the case of finite
types, clearly preserve entropy in the information-theoretic sense.

We will identify a ``featherweight'' version of $\Pi$ to use in our formal
development.  But first we illustrate the flavor of the family of languages with
some examples.

@@

>>>>

\usepackage{bbold}
\usepackage{bussproofs}
\usepackage{keystroke}
\usepackage{comment}
\usepackage{tikz}
\usepackage[inline]{enumitem}
\usepackage{agda}

@@

Starting from the physical principle of ``conservation of
information''~\cite{Hey:1999:FCE:304763,fredkin1982conservative}, James and
Sabry~\cite{James:2012:IE:2103656.2103667} propose a family of programming
languages $\Pi$ in which computation preserve information. Technically,
computations are \emph{type isomorphisms} which, at least in the case of finite
types, clearly preserve entropy in the information-theoretic
sense~\cite{James:2012:IE:2103656.2103667}. We will identify a ``featherweight''
version of $\Pi$ to use in our formal development.  But first we illustrate the
general flavor of the family of languages with some examples.

@@

\newcommand{\byiso}[1]{{\leftrightarrow}{\langle} ~#1~ \rangle}

{\small
\[\def\arraystretch{1.2}\begin{array}{rcl}
\AgdaFunction{toffoli} &:& \bt \otimes (\bt \otimes \bt)
                           \leftrightarrow  \bt \otimes (\bt \otimes \bt) \\
\AgdaFunction{toffoli} &=&
      \bt \otimes (\bt \otimes \bt) \\
&& \qquad\byiso{\AgdaFunction{unfoldBool} \otimes \AgdaFunction{id}} \\
&& (\ot \oplus \ot) \otimes (\bt \otimes \bt) \\
&& \qquad\byiso{\AgdaFunction{dist}} \\
&& (\ot \otimes (\bt \otimes \bt)) \oplus (\ot \otimes (\bt \otimes \bt)) \\
&& \qquad\byiso{\AgdaFunction{id} \oplus (\AgdaFunction{id} \otimes \AgdaFunction{cnot})} \\
&& (\ot \otimes (\bt \otimes \bt)) \oplus (\ot \otimes (\bt \otimes \bt)) \\
&& \qquad\byiso{\AgdaFunction{factor}} \\
&& (\ot \oplus \ot) \otimes (\bt \otimes \bt) \\
&& \qquad\byiso{\AgdaFunction{foldBool} \otimes \AgdaFunction{id}} \\
&& \bt \otimes (\bt \otimes \bt)
\end{array}\]}
